---
title: "Gapminder Data Analysis"
author: "Shawn Santo"
date: "2019-08-19"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      fig.width = 7)
```


## Load packages

## ctrl alt i inserts the coding chunk
```{r packages}
library(tidyverse)
```

## Load data

```{r data-read}
gap <- read_csv("https://bit.ly/gap_data")
```

## Exercises

### Exercise 1

Take a glimpse at the data set with function `glimpse()`.

```{r glimpse}
glimpse(gap)
```

How many variables and observations are in `gap`? What are the variable types
for the variables in `gap`?

Variable | Type
---------|----------
Country    | chr
Continent    | chr
Year    | dbl
Life Expectancy    | dbl
Population    | dbl
GDP per Capita    | dbl


### Exercise 2

#### Part 1

Create a plot of life expectancy versus year for the United States.

```{r usa}
gap_usa <- gap %>% 
  filter(country == "United States")

ggplot(data = gap_usa, mapping = aes(x = year, y = lifeExp)) +
  geom_point()
  
```

#### Part 2

Create a plot of life expectancy verses GDP per Capita for Canada. Set the
color of the points to be red.

```{r canada}
gap_ca <- gap %>% 
  filter(country == "Canada")

ggplot(data = gap_ca, mapping = aes(x = lifeExp, y = gdpPercap)) +
  geom_point(col="red") +
  geom_line()

```

#### Part 3

Create a plot of life expectancy versus GDP per Capita based on all countries
in Africa in 1992. Set the color to be purple.

```{r africa}
gap_a <- gap %>% 
  filter(continent == "Africa")

ggplot(data = gap_a, mapping = aes(x = lifeExp, y = gdpPercap)) +
  geom_point(col="purple")
```



### Exercise 3

#### Part 1

Create a plot of life expectancy versus GDP per Capita based on all the
countries in Africa and Europe in 1992. Set the point colors to code for the
two continents.

```{r africa-europe}
gap_EU_A <- gap %>% 
  filter(continent %in% c("Africa","Europe"),year == 1992)

ggplot(data = gap_EU_A, mapping = aes(x = lifeExp, y = gdpPercap,color = continent)) +
  geom_point()
```

#### Part 2

Create comparison boxplots by continent of GDP per Capita for the year 2007.
Fill the boxplots with color `darkgreen`.

```{r all-continents}
gap_07 <- gap %>% 
  filter(year == 2007)

ggplot(data = gap_07, mapping = aes(x = continent, y = gdpPercap)) +
  geom_boxplot(fill="darkgreen")
```


